manuscript_outline

### 2025-07-18 docs/manuscript/paper.Rmd
copy edit manuscript and rerender

Development in poikilothermic organisms is strongly temperature-dependent, particularly during early life stages prior to exogenous feeding. This thermally constrained physiology produces a near-mechanistic relationship between ambient temperature and developmental rate, allowing for predictive models of phenology. However, traditional models—often developed under constant laboratory temperatures—struggle to account for the variable thermal regimes experienced by wild populations. Effective value models were created to address this issue and are a framework that integrates daily temperature data to predict developmental timing under fluctuating conditions. Building on this foundation, the R package **hatchR** was developed for fishes to operationalize the effective value model and enable its widespread application. In this paper, we expand the scope of **hatchR**, demonstrating its applicability across a broad range of poikilothermic taxa including amphibians, reptiles, insects, crustaceans, copepods, and starfishes. We also illustrate how this framework supports both applied management goals—such as predicting hatching windows for conservation and monitoring—and basic research on ecological and evolutionary questions, including phenological plasticity, thermal adaptation, and ecological interactions. By providing a generalizable, open-source tool, **hatchR** bridges the gap between physiological theory and real-world application, offering a robust platform for predicting developmental phenology in diverse and changing environments.

@@

Because poikilothermic organisms rely on external sources to regulate their body temperature, the physiological processes governing their growth and development are tightly coupled to ambient thermal conditions. This relationship is especially deterministic during early life stages prior to exogenous feeding, when metabolic rates are primarily governed by environmental temperatures, modulated by genetic and maternal effects [@west-eberhard2003]. Although the absolute rate of development can vary widely among species and populations, most poikilotherms exhibit a consistent and often non-linear thermal performance curve that describes development as a function of temperature [@schoolfield1981; @angilletta2006]. These relationships can be effectively approximated with power law or exponential models that capture increasing developmental rates with rising temperature, up to a physiological optimum beyond which performance declines.

@@

To address these limitations in fishes, [@sparks2019] introduced the effective value model—an approach that captures the non-linear, temperature-dependent nature of development by integrating experimental data into a daily developmental rate function. Building on that work, the R package **hatchR** [@sparks2025] was developed to operationalize the effective value framework and make it accessible to researchers and managers. **hatchR** uses temperature-performance curves generated from lab-based thermal rearing trials to estimate the proportion of development achieved each day under observed temperature conditions. These daily contributions are summed until a cumulative threshold is reached (typically 1.0), marking the completion of a developmental stage. The package includes both an R-based workflow and a user-friendly Shiny web interface to facilitate broad adoption across research and management communities.

While hatchR was originally developed for coldwater fishes, the approach is broadly applicable to any poikilothermic organism whose development conforms to a monotonic or unimodal temperature–rate relationship, including many amphibians, reptiles, and invertebrates. In this manuscript, we expand the taxonomic scope of **hatchR**, compiling thermal developmental data from the literature and demonstrating how effective value modeling can be applied to diverse taxa. We also present three case studies illustrating how **hatchR** can inform ecological, evolutionary, and management questions—ranging from understanding thermal adaptation and phenological mismatches to forecasting hatch timing in conservation planning. These applications are increasingly relevant as ecosystems, especially aquatic ones, face intensifying and interacting stressors, including variable climates, habitat alteration, and invasive species, all of which can disrupt developmental timing and population dynamics [@craig2017; @pinsky2019].

@@

Effective value models estimate developmental progress based on empirically derived relationships between temperature and development rate. These relationships are typically parameterized by raising organisms at constant temperatures in laboratory settings, then fitting a non-linear function that captures the temperature-dependence of development. This function can then be reciprocated to produce daily “effective values”—units of development—that accumulate over time.

@@

**hatchR** requires two core inputs: a vector of daily average water temperatures and a corresponding vector of calendar dates. These data are typically collected from field-deployed temperature loggers. For loggers that record sub-daily temperatures, **hatchR** includes the function `summarize_temp()` to compute daily means. Data completeness and consistency can be checked using `plot_check_temp()` and `check_continuous()` to ensure there are no gaps or anomalies in the temperature record.

@@

If you import data from raw files with multiple daily readings, the package allows you to summarize your data with `summarize_temp()` and then check summarized data with the `plot_check_temp()` and `check_continuous()` functions.

@@

Developmental phenology can be predicted in two ways using hatchR. The `predict_phenology()` function estimates the date at which development completes, given a known reproductive event (e.g., oviposition or spawning; `spawn.date`) and daily temperature data. Alternatively, the `predict_spawn()` function works in reverse—estimating the likely date of a reproductive event based on a known or observed hatch/emergence date. This is especially useful for retrospective analyses or for estimating breeding windows from field observations.

@@

Examining variation in development across space and time is critical to understand how organisms respond to ecological factors such as variable climates [@pinsky2019]. Here, we first demonstrate an application of hatchR to predict hatching times from a spatio-temporally extensive dataset. Using these predictions, we then show how one can examine putative drivers of variation in development and discuss their application in forecasting future responses.

@@

Closely related species often diverge in their thermal tolerances and developmental strategies as part of niche differentiation. These differences can be visualized through reaction norms that capture genotype-by-environment (G×E) interactions, often derived from common garden or reciprocal transplant experiments. While such experiments are ideal for isolating genetic and plastic components of trait variation, they are logistically demanding and rarely available for many taxa.

@@

To demonstrate this approach, we leveraged data from @moore1939 to parameterize developmental models for four widely distributed North American frog species in the genus *Lithobates*: wood frog (*L. sylvaticus*), pickerel frog (*L. palustris*), northern leopard frog (*L. pipiens*), and green frog (*L. clamitans*). These species co-occur in parts of eastern North America but differ in their breeding phenology, habitat associations, and thermal niches. For example, wood frogs are early breeders in cold, ephemeral wetlands, while green frogs breed later in more permanent, warmer waters.

@@

As with interspecific comparisons, developmental models can be parameterized at the population level to examine local adaptation within species. In particular, hatchR allows researchers to compare how different populations respond to temperature, helping to reveal evolved differences in thermal sensitivity across environmental gradients.

Here, we used data from a common garden experiment on cabbage beetles (*Colaphellus bowringi*) reared at six constant temperatures (16–28 °C), spanning populations collected along a \~3,440 km latitudinal gradient in China [@tang2017]. Although all beetles were reared under identical conditions, developmental rates differed among populations, particularly at cooler temperatures. Populations from southern latitudes—where temperatures are generally warmer—developed more quickly than northern populations, a pattern consistent with cogradient variation, where genetic and environmental effects on a phenotype act in the same direction [@conover2009; @sparks2022].

To visualize these patterns, we fit custom effective value models for five of the six populations (excluding Longnan, which duplicated the values from Xiushui). The resulting linearized effective value curves (Figure 4) show that beetles from Xiushui County (southernmost) had consistently faster development rates across temperatures than their more northerly counterparts. These differences are especially pronounced at cooler temperatures, suggesting that local adaptation may only manifest under certain environmental conditions. Moreover, the consistent rank order of developmental rates across the populations indicates heritable differences among populations which have and can continue to be selected upon in changing environments.

@@

This example illustrates how **hatchR** can be used to explore the intersection of developmental plasticity and local adaptation using real-world data. Such applications are useful for disentangling the genetic basis of thermal sensitivity, anticipating geographic responses to changing temperatures, and informing population-specific management strategies.

@@

Although originally developed for fishes, **hatchR** is a generalizeable and flexible tool for predicting phenology across a broad diversity of poikilothermic organisms. The effective value framework accommodates nonlinear thermal sensitivity and allows researchers to estimate the timing of developmental milestones under realistic, fluctuating temperature regimes. In this manuscript, we demonstrate how power-law models can be fit using the `fit_model()` function, or defined externally and passed into the `predict_phenology()` function. For users without coding experience, a graphical user interface is also available via the **hatchR** Shiny app, expanding accessibility and encouraging broader adoption across disciplines.

We highlight the extensibility of **hatchR** through a compilation of parameterizable species across seven taxonomic classes and through three case studies that illustrate ecological, evolutionary, and conservation applications. These case studies span field data, phylogenetic comparisons, and experimental designs, demonstrating the tool’s flexibility across data types and biological questions.

@@

We also encourage users to carefully consider how models are parameterized. As discussed in greater detail in documentation for @sparks2025, we recommend using experimental data with at least four distinct temperature treatments spanning a wide thermal range. This is critical for capturing the nonlinear shape of developmental rate curves and avoiding overfitting or extrapolation errors. Many studies we encountered—particularly older or gray literature—relied on only two temperature treatments [*e.g.*, @qualls1998; @kozák2009]), which we do not recommend for model fitting due to the inability to detect curvature.

@@

We thank N. Karraker, D. Pilliod, M. Hayes, R.B. Bury, and B. Norman for their help in translating sites for the tailed frog example to locations that could be aligned with COMIDs. D. Horan helped organize GPS locations into COMIDs. The views expressed in this manuscript are those of the authors and do not necessarily represent the views or policies of USFS . Any mention of trade names, products, or services does not imply an endorsement by the U.S. government or USFS. USFS does not endorse any commercial products, services or enterprises.

@@

Morgan Sparks and Bryan Maitland developed the software and conceived the manuscript. All authors collected data and wrote and edited the manuscript. This manuscript describes an application for software for poikilotherms. As such, it has no specific locale for data collection, however significant effort was made to represent a globally and taxonomically diverse set of species for case studies and model sources.

@@

| Class | Order | Genus | Species | Study |
|:--------------|:--------------|:--------------|:--------------|:--------------|
| Amphibia | Anura | *Lithobates* | *L. sylvaticus, L. pipiens, L. clamitans, L. palustris* | @moore1939 |
|  |  | *Ascaphus* | *A. truei* | @brown1975 |
|  | Urodela | *Ambystoma* | *A. gracile* | @brown1976 |
| Reptilia | Squamata | *Sceloporus* | *S. undulatus* | @angilletta2000 |
|  |  | *Podarcis* | *P. muralis* | @vandamme1992 |
|  | Testudines | *Mauremys* | *M. reevesii* | @du2007 |
|  |  |  | 181 species | 141 studies in @while2018 |
| Insecta | Plecoptera | *Nemurella* | *N. pictetii* | @brittain1978, @elliott1984 |
|  |  | *Capnia* | *C. atra* | @brittain1984 |
|  |  | *Capnia* | *C. bifrons* | @elliott1986 |
|  |  | *Mesocapnia* | *M. oenone* | @brittain1984 |
|  |  | *Taeniopteryx* | *T. nebulosa* | @brittain1977 |
|  | Coleoptera | *Colaphellus* | *C. bowringi* | @tang2017 |
|  | Odonata | 7 genera | 14 species | Developmental equations in @pritchard1987 |
|  | Culicidae | *Aedes* | *A. sticticus, A. vexans* |  |
|  |  | Anopheles | *A. quadrimaculatus* |  |
|  |  | *Toxorhynchites* | *T. brevipalpis* |  |
| Malacostraca | Decapoda | *Pontastacus* | *P. leptodactylus* | @aydin2004 |
| Copepoda | Calanoida | 10 genera | 28 species | @forster2011 |
|  | Cyclopoida | *Limnoithona* | *L. tetraspina* |  |
|  |  | *Oithona* | *O. davisae* |  |
|  | Harpacticoida | *Microsetella* | *M. norvegica* |  |
|  |  | *Mesochra* | *M.lilljeborgi* |  |
|  | Poecilostomatoida | *Oncaea* | *O. venusta* |  |
| Cephalopoda | Octopoda | *Octopus* | *O. vulgaris, O. mimus* | @márquez2021 |
|  | Myopsida | *Loligo* | *L. vulgaris, L. reynaudii* |  |
|  | Oegopsida | *Illex* | *I. coindetii, I. illecebrosus* |  |
|  |  | *Todarodes* | *T.pacificus* |  |
|  |  | *Ommastrephes* | *O. bartramii* |  |
| Asteroidea | Valvatida | *Odontaster* | *O. meridionalis, O. validus* | @hoegh-guldberg1995 |
|  |  | *Asterina* | *A. miniata* |  |
|  |  | *Acanthaster* | *A. planci* |  |

@@

Figure 1: Custom hatching phenology model for coastal tailed frogs (*Ascaphus truei*).Panel A represents the model fit and raw data used to generate the effective value model and panel B are the effective values for daily temperatures between 6 and 20 °C. A dashed line with a 0.01 increase for every degree increase in included for reference.

@@

Figure 3: Custom hatch timing models developed for four North American *Lithobates* frog species. Models are parameterized from @moore1939 and phenology is predicted using two synthetic temperature regimes: Panel A shows predictions under a cooler regime (mean = 16 °C), while Panel B shows predictions under a warmer regime (mean = 24 °C). Panel C displays the effective value curves for each species—linearized approximations of thermal reaction norms. The crossover in rank order between Panels A and B illustrates genotype-by-environment interactions, reflecting species-specific thermal niches and breeding strategies.

>>>>

Development in poikilothermic organisms is strongly temperature-dependent, particularly during early life stages prior to exogenous feeding. This thermally constrained physiology produces a near-mechanistic relationship between ambient temperature and developmental rate, allowing for predictive models of phenology. However, traditional models—often developed under constant laboratory temperatures—struggle to account for the variable thermal regimes experienced by wild populations. Effective value models were created to address this issue and are a framework that integrates daily temperature data to predict developmental timing under fluctuating conditions. Building on this foundation, the R package **hatchR** was developed for fishes to operationalize the effective value model and enable its widespread application. In this paper, we expand the scope of **hatchR**, demonstrating its applicability across a broad range of poikilothermic taxa including amphibians, reptiles, insects, crustaceans, copepods, and starfishes. We also illustrate how this framework supports both applied management goals—such as predicting hatching windows for conservation and monitoring—and basic research on ecological and evolutionary questions including phenological plasticity, thermal adaptation, and ecological interactions. By providing a generalizable, open-source tool, **hatchR** bridges the gap between physiological theory and real-world application, offering a robust platform for predicting developmental phenology in diverse and changing environments.

@@

Because poikilothermic organisms rely on external sources to regulate their body temperature, the physiological processes governing their growth and development are tightly coupled to ambient thermal conditions. This relationship is especially deterministic during early life stages prior to exogenous feeding, when metabolic rates are primarily governed by environmental temperatures, though modulated by genetic and maternal effects [@west-eberhard2003]. Although the absolute rate of development can vary widely among species and populations, most poikilotherms exhibit a consistent and often non-linear thermal performance curve that describes development as a function of temperature [@schoolfield1981; @angilletta2006]. These relationships can be effectively approximated with power law or exponential models that capture increasing developmental rates with rising temperature, up to a physiological optimum beyond which performance declines.

@@

To address these limitations in fishes, @sparks2019 introduced the effective value model—an approach that captures the non-linear, temperature-dependent nature of development by integrating experimental data into a daily developmental rate function. Building on that work, the R package **hatchR** [@sparks2025] was developed to operationalize the effective value framework and make it accessible to researchers and managers. **hatchR** uses temperature-performance curves generated from lab-based thermal rearing trials to estimate the proportion of development achieved each day under observed temperature conditions. These daily contributions are summed until a cumulative threshold is reached (typically 1.0), marking the completion of a developmental stage. The package includes both an R-based workflow and a user-friendly Shiny web interface to facilitate broad adoption across research and management communities.

While**hatchR**was originally developed for coldwater fishes, the approach is broadly applicable to any poikilothermic organism whose development conforms to a monotonic or unimodal temperature–rate relationship, including many amphibians, reptiles, and invertebrates. In this manuscript, we expand the taxonomic scope of **hatchR**, compiling thermal developmental data from the literature and demonstrating how effective value modeling can be applied to diverse taxa. We also present three case studies illustrating how **hatchR** can inform ecological, evolutionary, and management questions—ranging from understanding thermal adaptation to forecasting hatch timing in conservation planning. These applications are increasingly relevant as ecosystems, especially aquatic ones, face intensifying and interacting stressors, including variable climates, habitat alteration, and invasive species, all of which can disrupt developmental timing and population dynamics [@craig2017; @pinsky2019].

@@

Effective value models estimate developmental progress based on empirically derived relationships between temperature and development rate. These relationships are typically parameterized by raising organisms at multiple constant temperatures in laboratory settings, then fitting a non-linear function that captures the temperature-dependence of development. This function can then be reciprocated to produce daily “effective values”—units of development—that accumulate over time.

@@

**hatchR** requires two core inputs: a vector of daily average temperatures and a corresponding vector of calendar dates. These data are typically collected from field-deployed temperature loggers. For loggers that record sub-daily temperatures, **hatchR** includes the function `summarize_temp()` to compute daily means. Data completeness and consistency can be checked using `plot_check_temp()` and `check_continuous()` to ensure there are no gaps or anomalies in the temperature record.

@@

@@

Developmental phenology can be predicted in two ways using hatchR. The `predict_phenology()` function estimates the date at which development completes, given a known reproductive event (*e.g.*, oviposition or spawning; using the `spawn.date` argument) and daily temperature data. Alternatively, the `predict_spawn()` function works in reverse—estimating the likely date of a reproductive event based on a known or observed hatch/emergence date. This is especially useful for retrospective analyses or for estimating breeding windows from field observations.

@@

Examining variation in development across space and time is critical to understand how organisms respond to ecological factors such as variable climates [@pinsky2019]. Here, we demonstrate an application of **hatchR** to predict hatching times from a spatio-temporally extensive dataset. Using these predictions, we then show how one can examine putative drivers of variation in development and discuss their application in forecasting future responses.

@@

Closely related species often diverge in their thermal tolerances and developmental strategies as part of niche differentiation [@angilletta2009]. These differences can be visualized through reaction norms that capture genotype-by-environment (G×E) interactions, often derived from common garden or reciprocal transplant experiments. While such experiments are ideal for isolating genetic and plastic components of trait variation, they are logistically demanding and rarely available for many taxa.

@@

To demonstrate this approach, we leveraged data from @moore1939 to parameterize developmental models for four widely distributed North American frog species in the genus *Lithobates*: wood frog (*L. sylvaticus*), pickerel frog (*L. palustris*), northern leopard frog (*L. pipiens*), and green frog (*L. clamitans*). These species co-occur in parts of eastern North America but differ in their breeding phenology, habitat associations, and thermal niches. For example, wood frogs are early breeders in cold, ephemeral wetlands, while green frogs breed later in more permanent, warmer waters [@moore1939].

@@

As with interspecific comparisons, developmental models can be parameterized at the population level to examine local adaptation within species. In particular, **hatchR** allows researchers to compare how different populations respond to temperature, helping to reveal evolved differences in thermal sensitivity across environmental gradients.

Here, we used data from a common garden experiment on cabbage beetles (*Colaphellus bowringi*, females) reared at six constant temperatures (16–28 °C), spanning populations collected along a \~3,440 km latitudinal gradient in China [@tang2017]. Although all beetles were reared under identical conditions, developmental rates differed among populations, particularly at cooler temperatures. To visualize these patterns, we fit custom effective value models for five of the six populations (excluding Longnan, which duplicated the values from Xiushui). The resulting linearized effective value curves (Figure 4) show that beetles from Xiushui County (southernmost) had consistently faster development rates across temperatures than their more northerly counterparts, a pattern consistent with cogradient variation, where genetic and environmental effects on a phenotype act in the same direction [@conover2009; @sparks2022]. These differences are especially pronounced at cooler temperatures, suggesting that local adaptation may only manifest under certain environmental conditions. Moreover, the consistent rank order of developmental rates across the populations indicates heritable differences among populations which have and can continue to be selected upon in changing environments.

@@

This example illustrates how **hatchR** can be used to explore the intersection of developmental plasticity and local adaptation in a novel framework. Such applications are useful for disentangling the genetic basis of thermal sensitivity, anticipating geographic responses to changing temperatures, and informing population-specific management strategies.

@@

Although originally developed for fishes, **hatchR** is a generalizeable and flexible tool for predicting phenology across a broad set of poikilothermic organisms. The effective value framework accommodates nonlinear thermal sensitivity and allows researchers to estimate the timing of developmental milestones under real-world, fluctuating temperature regimes with high accuracy and precision. In this manuscript, we demonstrate how power-law models can be fit using the `fit_model()` function, or defined externally and passed into the `predict_phenology()` function. For users without coding experience, a graphical user interface is also available via the **hatchR** Shiny app, expanding accessibility and encouraging broader adoption across disciplines.

We highlight the extensibility of **hatchR** through a compilation of parameterizable datasets spanningseven taxonomic classes and through three case studies that illustrate ecological, evolutionary, and conservation applications. These case studies span field data, phylogenetic comparisons, and experimental designs, demonstrating the tool’s flexibility across data types and biological questions.

@@

We also encourage users to carefully consider how models are parameterized. As discussed in greater detail in documentation for @sparks2025, we recommend using experimental data with at least four distinct temperature treatments spanning a wide thermal range. This is critical for capturing the nonlinear shape of developmental rate curves and avoiding overfitting or extrapolation errors. Many studies we encountered—particularly older or gray literature—relied on only two temperature treatments [*e.g.*, @qualls1998; @kozák2009], which we do not recommend for model fitting due to the inability to detect curvature.

@@

We thank N. Karraker, D. Pilliod, M. Hayes, R.B. Bury, and B. Norman for their help in translating sites for the tailed frog example to locations that could be aligned with COMIDs. D. Horan helped organize GPS locations into COMIDs. The views expressed in this manuscript are those of the authors and do not necessarily represent the views or policies of USFS. Any mention of trade names, products, or services does not imply an endorsement by the U.S. government or USFS. USFS does not endorse any commercial products, services or enterprises.

@@

Morgan Sparks and Bryan Maitland developed the software and conceived the manuscript. All authors collected data sources and wrote and edited the manuscript.

@@

| Class | Order | Genus | Species | Study |
|:--------------|:--------------|:--------------|:--------------|:--------------|
| Amphibia | Anura | *Lithobates* | *L. sylvaticus, L. pipiens, L. clamitans, L. palustris* | @moore1939 |
|  |  | *Ascaphus* | *A. truei* | @brown1975 |
|  | Urodela | *Ambystoma* | *A. gracile* | @brown1976 |
| Reptilia | Squamata | *Sceloporus* | *S. undulatus* | @angilletta2000 |
|  |  | *Podarcis* | *P. muralis* | @vandamme1992 |
|  | Testudines | *Mauremys* | *M. reevesii* | @du2007 |
|  |  |  | 181 species | 141 studies in @while2018 |
| Insecta | Plecoptera | *Nemurella* | *N. pictetii* | @brittain1978, @elliott1984 |
|  |  | *Capnia* | *C. atra* | @brittain1984 |
|  |  | *Capnia* | *C. bifrons* | @elliott1986 |
|  |  | *Mesocapnia* | *M. oenone* | @brittain1984 |
|  |  | *Taeniopteryx* | *T. nebulosa* | @brittain1977 |
|  | Coleoptera | *Colaphellus* | *C. bowringi* | @tang2017 |
|  | Odonata | 7 genera | 14 species | Developmental equations in @pritchard1987 |
|  | Culicidae | *Aedes* | *A. sticticus, A. vexans* |  |
|  |  | *Anopheles* | *A. quadrimaculatus* |  |
|  |  | *Toxorhynchites* | *T. brevipalpis* |  |
| Malacostraca | Decapoda | *Pontastacus* | *P. leptodactylus* | @aydin2004 |
| Copepoda | Calanoida | 10 genera | 28 species | @forster2011 |
|  | Cyclopoida | *Limnoithona* | *L. tetraspina* |  |
|  |  | *Oithona* | *O. davisae* |  |
|  | Harpacticoida | *Microsetella* | *M. norvegica* |  |
|  |  | *Mesochra* | *M.lilljeborgi* |  |
|  | Poecilostomatoida | *Oncaea* | *O. venusta* |  |
| Cephalopoda | Octopoda | *Octopus* | *O. vulgaris, O. mimus* | @márquez2021 |
|  | Myopsida | *Loligo* | *L. vulgaris, L. reynaudii* |  |
|  | Oegopsida | *Illex* | *I. coindetii, I. illecebrosus* |  |
|  |  | *Todarodes* | *T.pacificus* |  |
|  |  | *Ommastrephes* | *O. bartramii* |  |
| Asteroidea | Valvatida | *Odontaster* | *O. meridionalis, O. validus* | @hoegh-guldberg1995 |
|  |  | *Asterina* | *A. miniata* |  |
|  |  | *Acanthaster* | *A. planci* |  |

@@

Figure 1: Custom hatching phenology model for coastal tailed frogs (*Ascaphus truei*). Panel A represents the model fit and raw data used to generate the effective value model and panel B are the effective values for daily temperatures between 6 and 20 °C. A dashed line with a 0.01 increase for every degree increase in included for reference.

@@

Figure 3: Custom hatch timing models developed for four North American *Lithobates* frog species. Models are parameterized from @moore1939 and phenology is predicted using two synthetic temperature regimes: Panel A shows predictions under a cooler regime (mean = 16 °C), while Panel B shows predictions under a warmer regime (mean = 24 °C). Panel C displays the effective value curves for each species—linearized thermal reaction norms. The crossover in rank order between Panels A and B illustrates genotype-by-environment interactions, reflecting species-specific thermal niches and breeding strategies.
